---
title: "Meta-analysis of the rodent object-in-context task: Random forest-based Meta-analysis (MetaForest)"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

# Prepare environment
```{r setup, include=FALSE}
rm(list = ls()) #clean environment
# packages
library(dplyr)
library(caret) #for cross-validation
library(metaforest) #for exploratory analysis
# to check which packages are loaded
# (.packages())
```

# Metaforest
https://www.rdocumentation.org/packages/metaforest/versions/0.1.3

## Data preparation
```{r metaforest preparation}
readRDS("processed_data/data_with_effect_size.RDS") ->data6
str(data6)

# create dataset for MetaForest
data6 %>% 
  select( -c(PMID, EXP_group, 
             SmellCntrl, # only 1 level, so non-informative
             nC,
             # other outcome variables
             DR, DR_A, DR_B, SEM, SEM_DR_A, SEM_DR_B, SD, label,
             no.Affiliation, # captures mostly the same variation as `PMID` variable (no.Affiliation: 29 levels, PMID: 37 levels)
             Program # captures mostly the same variation as in the 'scoring' variable.
  )) %>% 
  droplevels() -> datForest
is.na(datForest) %>% any()
```

## Save data for WS follow-up plots
```{r save data WS plots}
data6 %>% select(nC, ID, PMID)->sample.sizes
full_join(sample.sizes, datForest, by="ID")->datForest_for_WS_Plot
saveRDS(datForest_for_WS_Plot, "processed_data/datForest_for_WS_Plot.RDS")
```

## MetaForest: Tuning
Note, tuning assumes that convergence is reached. To set-up metaforest, also see: https://www.rdocumentation.org/packages/metaforest/versions/0.1.3/topics/ModelInfo_mf
```{r metaforest tuning}
set.seed(94372)
ModelInfo_mf()$notes # information on method from MetaForest
# Set up 9-fold grouped cross-validation
fit_control <- trainControl(method = "cv", number = 9, index = groupKFold(datForest$ID, k = 9)) 
# Set up a custom tuning grid for the three tuning parameters of MetaForest
rf_grid <- expand.grid(whichweights = c("random", "fixed", "unif"),
                       mtry = c(2, 4, 6),
                       min.node.size = c(2, 4, 6))
```

## MetaForest: Train Model
Cross-validated clustered MetaForest
```{r MetaForest training}
set.seed(94372)
# Train the model
cv.mf.cluster <- train(y = datForest$yi, x = datForest[names(datForest) !="yi"],
                       study = "ID", 
                       method = ModelInfo_mf(), #= the method from MetaForest. (https://rdrr.io/cran/metaforest/man/MetaForest.html)
                       trControl = fit_control,
                       tuneGrid = rf_grid,
                       num.trees = 600,
                       data=datForest)
# Save the model
saveRDS(cv.mf.cluster,file="processed_data/fitted.MetaForest.RDS")
```

To load saved model directly
```{r eval=FALSE, include=FALSE}
readRDS("processed_data/fitted.MetaForest.RDS")->cv.mf.cluster
```

## MetaForest: Inspect model

### Tuning parameters
#### Show model: The final values used for the model were whichweights = unif, mtry = 4 and min.node.size = 2
```{r inspect model 1}
cv.mf.cluster
```
#### Plot grid parameter models
```{r inspect model 2}
plot(cv.mf.cluster)
```

### Model Summary
```{r inspect model 3}
summary(cv.mf.cluster)
```

### R(oob)
https://towardsdatascience.com/what-is-out-of-bag-oob-score-in-random-forest-a7fa23d710
```{r inspect model 4}
cv.mf.cluster$finalModel
```

### Cross validated R2 with SD
details for the "best" model: unif, mtry = 4 and min.node.size = 2
```{r inspect model 5}
cv.mf.cluster$results[which(cv.mf.cluster$results$whichweights == "unif" & 
                              cv.mf.cluster$results$mtry == 4 & 
                              cv.mf.cluster$results$min.node.size == 2),] 
```

### Convergence plot
Good model convergence
```{r convergence plot}
jpeg(file="results/metaforest_convergencePlot.jpeg", height = 3000, width = 3000, res=500)
plot(cv.mf.cluster$finalModel) 
dev.off()
```

## Variable importance (based on metaForest)
```{r variable importance plot}
tiff(file="results/metaforest_varImportance.tiff", height=1500, width=1500, pointsize=8, res=300)
VarImpPlot(cv.mf.cluster$finalModel)+
  geom_hline(yintercept=11.5) +
  annotate(geom="text", x=0.0017, y=12.5, label="Top 50%", color="darkblue")
dev.off()
```

## Export important variables
```{r metaforest varimportance scores}
varImp(cv.mf.cluster) ->imp.scores

# Export variable importance scores (ordered) to csv
imp <- as.data.frame(imp.scores$importance)
imp <- data.frame(overall = imp$Overall,
                  names   = rownames(imp))
write.csv2(imp[order(imp$overall,decreasing = T),],"results/important_variables_metaforest.csv")

# Select top 50% most important variables
imp[order(imp$overall,decreasing = T),] %>%
  slice_head(.,n=nrow(.) / 2) -> important.vars 
important.vars$names ->variable
variable

# Save top 50% most important variables
saveRDS(variable,"processed_data/important_variables.RDS")
```