---
title: "Meta-analysis of the rodent object-in-context task: prepare data"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
editor_options: 
  chunk_output_type: console
---

# prepare environment
```{r setup, include=FALSE}
rm(list = ls()) #clean environment
# packages
library(readxl)
library(dplyr)
library(osfr)
```

# download from OSF & load data
```{r}
osf_retrieve_file("qadch") %>% osf_download(path = "data", conflicts="overwrite")
# load data
read_excel("data/280121_Data_Extraction_RoB.xlsx", sheet = "Extraction_converted", na=c("N/A", "#VALUE!"))->data
#  check empty rows?
which(rowSums(is.na(data) ) == ncol(data))
```

# Rename variables
```{r rename}
data %>% rename(
  DR_B=`DR_B_n/(f+n)`, 
  DR_A=`DR_A_(n-f)/(n+f)`,
  Prev.Exp = PrevExp,
  Prev.OR = Prev_OR,
  Type.Control = TypeCntrl,
  Context.Floor = C_Floor,
  Context.Wall = C_Walls,
  Context.Room = C_Rooms,
  Context.Size.A = Box_Size_A,
  Context.Size.B = Box_Size_B,
  Context.Shape = B_shapes,
  Object.Material = O_Mat,
  Object.Size = O_Size,
  Object.Position = O_Position,
  Context.Habituation.Freq = H_BoxesFreq,
  Object.Habituation = H_Object,
  Testroom.Habituation = H_Testroom,
  Transport.Habituation = H_Transport,
  Habituation.with.Cagemates = H_mates,
  Habituation.Time.Trial = `H_Time(min)_per_trial`,
  Habituation.Time.Total.Context = `H_Time_Total_per_Box`,
  
  # Values below will be recalculated after missing are corrected (see section "recreate sumscores")
  # Arousal.Prior = Arousal_Score,
  # Context.Difference.Score = Context_Score,
  # Arousal.Task.Habituation = Hab_Arousal_Score,
  # Arousal.Total = Total_Arousal_Score, # idem
  
  Learning.Time.Trial = L_Time_per_box,
  Learning.Learning.Delay = L_retention,
  Learning.Order.Counterbalanced = L_counter,
  Learning.Exploration.Time.A = L_Expl_boxA,
  Learning.Exploration.Time.B = L_Expl_boxB,
  Learning.Exploration.Time.Total = L_Expl_tot,
  Learning.Memory.Delay = Retention,
  Memory.Time.Trial = M_Time,
  Memory.Exploration.Time.A = M_Expl_boxA,
  Memory.Exploration.Time.B = M_Expl_boxB,
  Memory.Exploration.Time.Familiar = M_In_ctxt,
  Memory.Exploration.Time.New = M_Out_ctxt,
  Memory.Exploration.Time.Total = M_Expl_tot,
  Context.Order = C_order,
  DR.Formula=Formula,
  Memory.Type = `Type_Mem (according to article)`)->df
# str(df)
```

- DR_A center is at 0    [*2-1]
- DR_B center is at 0.5   [*0.5+0.5]

# Create new variables
```{r rename}
df %>%
  mutate(
    # create unique id
    ID=paste(PMID, EXP_group,sep="_"),
    # create first author column
    Author=gsub("([A-Za-z]+).*", "\\1", Authors)
  ) %>%
  mutate(
    #create label for graph
    label=paste(Author, Year,sep=" "),
    # recode scoring program var
    Program = case_when(
      grepl("Any", Program) ~ "AnyMaze software",
      grepl("n.a.", Program) ~ "n.a.",
      TRUE ~ Program) # "catch all" test see: https://stackoverflow.com/questions/39257820/keep-value-if-not-in-case-when-statement
  )->data1  
str(data1)
```

# Save data for systematic review
```{r}
saveRDS(data1, "processed_data/SR_data.RDS")
```

# Missing values

## Exclude studies with insufficient data for meta-analysis
Note, these studies will be included in systematic review only
```{r}
summary(data1)
# Select studies with missing values in 'meta-analysis' variables.
data1 %>% 
  filter_at(vars(c("nC", "DR", "SEM", "DR_A","DR_B","SEM_DR_A", "SEM_DR_B")), any_vars(is.na(.))) %>%
  select(PMID) %>%
  unique()
# 4 studies missing values are excluded from meta-analysis, but included in systematic review

data1 %>% 
  filter_at(vars(c("nC", "DR", "SEM", "DR_A","DR_B","SEM_DR_A", "SEM_DR_B")), all_vars(!is.na(.))) %>%
  select(PMID) %>%
  unique()
# 37 studies included in meta-analysis

data1 %>% 
  filter_at(vars(c("nC", "DR", "SEM", "DR_A","DR_B","SEM_DR_A", "SEM_DR_B")), all_vars(!is.na(.))) -> data_meta
```

## Exclude variables that have 1/3 missing values, or are non-informative for the meta-analysis
```{r missing}
# variables with more than 1/3 missing values
which(colSums(is.na(data_meta)) > (1/3)*nrow(data_meta)) -> too.much.missing
# save excluded variables
names(too.much.missing) %>% write.csv2("results/too.much.missing.vars.csv")

# non-informative variables for meta-analysis
Variables.to.exclude <- c("Reference (APA)",
                          "Data_Extracted",
                          "Authors", "Affiliation",
                          "Journal", 
                          "Author",
                          "Page_outcome", "Page_outcome", "Def_Expl")

data_meta %>% 
  select(-all_of(too.much.missing)) %>%
  select(-all_of(Variables.to.exclude)) %>% 
  droplevels()->data_meta2
# str(data2)
```

## Recode factor variable types
```{r recode factors}
factors<- c("ID", "PMID", "EXP_group", "no.Affiliation", "Species", "Strain", "Sex", "Age", "DR.Formula", "Memory.Type", "Scoring", "Object.Habituation", "Object.Material", "Context.Shape", "Context.Wall", "Context.Floor", "RevRhythm", "Housing", "AdLib", "Type.Control", "Prev.OR", "Prev.Exp", "Origin", "Habituation.with.Cagemates", "Program", "Learning.Order.Counterbalanced","Context.Order", "SmellCntrl")
data_meta2 %>% 
  mutate_at(., factors, as.character) %>% 
  mutate_at(., factors, factor) -> data_meta3 
# str(data_meta3)
```

## Fix missing values (in variables with less then 1/3 missing): 
Missing values are replaced by median value (for numeric) of most prevalent category (for factors)
NOTE. Strain (3 missing values) is dependent upon Species, missing values are replaced by 'species-specific' median.
```{r fix missing values}
# check missing values
summary(data_meta3)

factors.to.fix <- c("Sex", "Origin", "Age", "Prev.Exp", "Prev.OR", "AdLib", "Housing", "RevRhythm", "Context.Floor", "Context.Wall", "Context.Shape","Context.Room", "Habituation.with.Cagemates", "Program", "Object.Material", "Object.Habituation", "Learning.Order.Counterbalanced", "Context.Order", "Scoring", "SmellCntrl", "Memory.Type")

numeric.to.fix <- c( "Context.Size.A", "Context.Size.B","Context.Habituation.Freq", "Habituation.Time.Trial", "Habituation.Time.Total.Context", "Learning.Learning.Delay", "Memory.Time.Trial", "Learning.Memory.Delay")

# change character to numeric
data_meta3$Habituation.Time.Trial<-as.numeric(data_meta3$Habituation.Time.Trial)

for (i in c(1:length(numeric.to.fix))) {
  numeric.to.fix[[i]] -> varname
  data_meta3[is.na(data_meta3[,varname]),varname] <- median(unlist(data_meta3[,varname]), na.rm = T)
}

# categorical missing: substitute with most common category
categorical_mode <- function(x){
  names(table(x))[which.max(table(x))]
}

for (i in c(1:length(factors.to.fix))) {
  factors.to.fix[[i]] -> varname
  data_meta3[is.na(data_meta3[,varname]),varname] <- categorical_mode(unlist(data_meta3[,varname]))
}

# for checking
summary(data_meta3)
any(is.na(data_meta3))


# define median mice
data_meta3 %>% filter(Species == "Mice") %>% select("Strain") %>% unlist() %>% categorical_mode() -> median.mice
# Define median rat
data_meta3 %>% filter(Species == "Rat") %>% select("Strain") %>% unlist() %>% categorical_mode() -> median.rat

# for checking
data_meta3 %>% filter (is.na(Strain) & (Species == "Rat") ) #1
data_meta3 %>% filter (is.na(Strain) & (Species == "Mice") ) #2

# replace missing values with species specific median
data_meta3 %>% mutate(
  Strain = case_when(
    # !is.na(Strain) ~ Strain,
    is.na(Strain) & (Species == "Rat") ~ median.rat,
    is.na(Strain) & (Species == "Mice") ~ median.mice,
    TRUE ~ as.character(Strain))
) -> data_meta4

# for checking
any(is.na(data_meta4))

# count missing without meta-analysis variables and no.affiliation
data_meta4 %>% 
  select(-c(no.Affiliation, nC, DR, SEM, DR_A, DR_B, SEM_DR_A, SEM_DR_B)) %>% #View()
  is.na(.) %>% #View()
  any() # no missing values

summary(data_meta4)
```

# Recreate sumscores (after missing corrections)
The original object score was created by  "Object_Score = Object.Material2 + Object.Size2". But Object.Size is removed from analysis dataset, because there were too much missing values. Note, there is no object sum score created, as this would be identical to Object.Material.

The variables  Handling, Transport.Habituation2, Transport.Habituation, Testroom.Habituation were also excluded because there were too much missing values. These variable were therefor not included 
```{r}
summary(data_meta4)

data_meta4 %>%
  mutate(
    # for arousal score
    Origin2 = case_when(
      Origin == "B" ~ 0,
      Origin == "P" ~ 1),
    Prev.Exp2 = case_when(
      Prev.Exp == "No" ~ 0,
      Prev.Exp == "Yes" ~ 1),
    Prev.OR2 = case_when(
      Prev.OR == "No" ~ 1,
      Prev.OR == "Yes" ~ 0),
    Type.Control2 = case_when(
      Type.Control == "U" ~ 0,
      Type.Control == "SI" ~ 1,
      Type.Control == "SO" ~ 2,),
    AdLib2 = case_when(
      AdLib == "Yes" ~ 0,
      AdLib == "No" ~ 1),
    RevRhythm2 = case_when(
      RevRhythm == "Yes" ~ 0,
      RevRhythm == "No" ~ 1),
    Housing2 = case_when(
      Housing == "G" ~ 0,
      Housing == "S" ~ 1),
    Arousal.Prior = Origin2 + Prev.Exp2 + Prev.OR2 + Type.Control2 + AdLib2 + RevRhythm2 + Housing2,
    
    # Context.Difference.Score
    Context.Floor2 = case_when(
      Context.Floor == "Different" ~ 4,
      Context.Floor == "Similar" ~ 0), # if no match, case_when() creates NA
    Context.Wall2 = case_when(
      Context.Wall == "Different" ~ 3,
      Context.Wall == "Similar" ~ 0),
    Context.Room2 = case_when(
      Context.Room == "Different" ~ 1,
      Context.Room == "Similar" ~ 0),
    Context.Shape2 = case_when(
      Context.Shape == "Different" ~ 1,
      Context.Shape == "Similar" ~ 0),
    Context.Difference.Score = Context.Floor2 + Context.Wall2 + Context.Room2 + Context.Shape2,
    
    # Object score (excluded, too much missing values)
    # Object.Material2 = case_when(
    #   Object.Material == "Different" ~ 1,
    #   Object.Material == "Similar" ~ 0),
    # Object.Size2 = case_when(
    #   Object.Size == "Different" ~ 1,
    #   Object.Size == "Similar" ~ 0),
    # Object_Score = Object.Material2 + Object.Size2,
    
    # Arousal.Task.Habituation (commented variables were excluded from data, too much missing values)
    # Handling2 = case_when(
    #   Handling == "Yes" ~ 0,
    #   Handling == "No" ~ 1),
    # Testroom.Habituation2 = case_when(
    #   Testroom.Habituation == "Yes" ~ 0,
    #   Testroom.Habituation == "No" ~ 1),
    # Transport.Habituation2 = case_when(
    #   Transport.Habituation == "Yes" ~ 0,
    #   Transport.Habituation == "No" ~ 1),
    Context.Habituation.Freq2 = case_when(
      Context.Habituation.Freq == 0 ~ 1,
      Context.Habituation.Freq > 0 ~ 0),             
    Habituation.with.Cagemates2 = case_when(
      Habituation.with.Cagemates == "Yes" ~ 0,
      Habituation.with.Cagemates == "Partly" ~ 0,
      Habituation.with.Cagemates == "n.a." ~ 0, # Note: n.a scored when, for example, when no habituation was performed
      Habituation.with.Cagemates == "No" ~ 1),
    # Transport.Habituation2 = case_when(
    #   Transport.Habituation == "Yes" ~ 0,
    #   Transport.Habituation == "Partly" ~ 0,
    #   Transport.Habituation == "No" ~ 1),   
    Object.Habituation2 = case_when(
      Object.Habituation == "Yes" ~ 0,
      Object.Habituation == "Partly" ~ 0,
      Object.Habituation == "n.a." ~ 0, # Note: n.a scored when, for example, when no habituation was performed
      Object.Habituation == "No" ~ 1 ), 
    Arousal.Task.Habituation = # Handling2 + Testroom.Habituation2 + Transport.Habituation2 + #Transport.Habituation2+
      Context.Habituation.Freq2 + Habituation.with.Cagemates2 + Object.Habituation2, # higher score more arousal
    
    # Total arousal        
    Arousal.Total = Arousal.Prior + Arousal.Task.Habituation)  %>%
  # remove helper/intermediate variables (created above)
  select(-c(Context.Habituation.Freq2, Object.Habituation2,Context.Habituation.Freq2, Habituation.with.Cagemates2, Context.Floor2, Context.Wall2, Context.Room2, Context.Shape2, Origin2 , Prev.Exp2 , Prev.OR2, Type.Control2 , AdLib2 , RevRhythm2, Housing2)
  )-> data_meta5
```

# Change order and names of factor levels
For readability of the MetaForest follow-up plots
```{r}
data_meta5 %>%
  mutate(
    Memory.Type = factor(Memory.Type, 
                         levels=c("STM", 'medium', "LTM")),
    DR.Formula = factor(DR.Formula, 
                        levels = c("(tnew-tfamiliar)/(tnew+tfamiliar)", "[(tnew-tfamiliar)/(tnew+tfamiliar)]*100",
                                   "(tnew)/(tnew+tfamiliar)",  "[(tnew)/(tnew+tfamiliar)]*100"),
                        labels = c("n-f/t", "n-f/t*100", 
                                   "n/t", "n/t*100")),
    Strain = factor(Strain, 
                    levels = c(
                      # mice
                      "Tg(Sim1cre)KH21Gsat/Mmucd", "Sv/Ev mice", "ICR", "C57BL/6",
                      # rats
                      "Lister hooded rats", "Long-Evans rats", "Sprague-Dawley", "Wistar rats", "pigmented DA strain", "Dark Agouti rats"),
                    labels = c(
                      # mice
                      "TMm", "SEm", "ICm", "BLm",
                      # rats
                      "LHr", "LEr", "SDr", "Wr", "pDr", "DAr")
                    ),
    Age = factor(Age, 
                 levels = c("Juvenile", "Adolescent", "Adult", "Senior")),
    Type.Control = factor(Type.Control, 
                          levels = c("U", "SI", "SO")),
    Program = factor(Program, 
                     levels = c("AnyMaze software", "Ethovision XT (Noldus)", 
                                "Med-PCIV software", "Topscan", "Limelight2 program (Actimetrics; Coubourn Instruments)", "n.a."),
                     labels = c("Any", "Eth",
                                "Med", "Top", "LLP", "non"))) %>% 
  droplevels() ->data_meta6
```

# Save processed data
```{r}
saveRDS(data_meta6, "processed_data/cleaned_data.RDS")
```